---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# geolinkr

geolinkr is an R package that creates crosswalks for empirical comparisons across geographies and over time using shapefiles. 

You can create a crosswalk by providing three shapefiles to the `create_cw()` function:

- `from_sf`: A shapefile with the **source** geographies as an [sf object](https://github.com/r-spatial/sf?tab=readme-ov-file#simple-features-for-r) 
- `to_sf`: A shapefile with the **target** geographies as an sf object.
- `wts_sf`: A shapefile usually with census blocks or tracts delineations and a weighting variable, such as population or household counts, to allocate geographies from `from_sf` to `to_sf`. 

See [Examples](#Examples)
	

## Installation

``` r
# install.packages("remotes")
remotes::install_github("ChandlerLutz/geolinkr")
```

## Examples

In 2023, Connecticut updated its county definitions, creating a break in various economic datasets. We can use the `geolinkr::create_cw()` function to create a crosswalk from the Connecticut 2020 to the 2023 county definitions, using household counts at the tract level as weights.

- Note: It's typically best to use the smallest available delineations, such as Census blocks, for the weights, but we'll use tracts to keep the example tractable.

First, download the 2020 CT county delineations (the source) and the 2023 CT county delineations (the target):

```{r dwnld, warning=FALSE, message=FALSE, results='hide'}
library(geolinkr)
library(sf)
library(tigris)

# 2020 CT counties -- source shapefile 
ct_cnty20 <- tigris::counties(state = "CT", year = 2020) |>
  sf::st_transform(crs = 5070) |>
  _[, c("GEOID", "geometry")]
names(ct_cnty20) <- c("geoid", "geometry")

# 2023 CT counties -- target shapefile
ct_cnty23 <- tigris::counties(state = "CT", year = 2023) |>
  sf::st_transform(crs = 5070) |>
  _[, c("GEOID", "geometry")]
names(ct_cnty23) <- c("geoid", "geometry")
```

From 2020 to 2023, Connecticut's county count increased from 8 to 9 and the definitions changed:

```{r ct_cnty_chg}
par(mfrow = c(1, 2))
plot(ct_cnty20$geometry, main = "2020 CT counties")
plot(ct_cnty23$geometry, main = "2023 CT counties")
```

We'll use the number of housing units in each census tract as weights. I have prepared other tract, block group, and block shapefiles dating back to 2000 that you can use for weights [here](https://github.com/ChandlerLutz/census-blocks-tracks-shp).

```{r wts}
# 2020 CT census tracts -- wts shapefile (with hh2020 as the weight variable)
ct_tracts20 <- readRDS(url("https://github.com/ChandlerLutz/ct-shps/blob/main/ct_tracts20_sf.rds?raw=1")) |>
  sf::st_transform(crs = 5070) |>
  _[, c("GEOID", "hh2020", "geometry")]
names(ct_tracts20) <- c("geoid", "hh2020", "geometry")

print(ct_tracts20)
```

Finally, we'll create the crosswalk using the `create_cw()` function. The output is the crosswalk from 2020 CT counties to 2023 CT counties.

```{r create_cw}
cw_ct_cnty20_cnty23 <- create_cw(
  from_sf = ct_cnty20,
  to_sf = ct_cnty23,
  wts_sf = ct_tracts20,
  wt_var_name = "hh2020",
  check_that_wts_cover_from_and_to = FALSE
)

print(cw_ct_cnty20_cnty23)
```
